---
title: "VVnA R Package"
author: "Aous Abdo"
date: "Tuesday, August 12, 2014"
output:
  pdf_document:
    toc: TRUE
    toc_depth: 3
    number_sections: true
---
```{r, echo=FALSE,  warning=FALSE}
library(knitr)
library(VVnA)
opts_chunk$set(eval=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=10))
```
# Introduction
 
The VVnA "Validation, Verification, and Accreditation" package is a package intended 
for ....

# Air Resistance
Air resistance is a friction force resulting from different atmoic phenomena and depends on the velocity of the projectile relative to air. For most objects, the direction of  the force is always opposite to the main velocity vector.  

## Air Resistance at Low Speeds
At low speeds, the force of friction with air can be approximated as a summation 
of linear an duadratic terms:  
$$
\begin{aligned}
f(v) & = f_{lin} + f_{quad} \\
& = bv + cv^2
\end{aligned}
$$

For a spherical object of diametere $D$, the coefficients of friction are: 

$$
\begin{aligned}
b & = \beta D \\
c & = \gamma D^2
\end{aligned}
$$
Where 
$$
\begin{aligned}
\beta &  = 1.6 \times 10^{-4} \, N.s/m^2 \\
\gamma & = 0.25 \, N.s/m^4
\end{aligned}
$$

For more details on the derivation of the equations used in this package please 
see this [document](http://rhig.physics.wayne.edu/~pruneau/Courses/PHY5200/lectures/PHY5200-Chap2.pdf)

# Functions
## Projectile Motion
Projectile motion in vacuum and in air are calculated with the `projectile` and `projFrictionLin` functions respectively. when considering air friction effects on projectiles, we only consider the viscous drag which is related to the 
velocity `v`. The Inertial drag realted to the square of the velocity is not treated in this package. 

In each of these two cases, a function will return the following projectile parameters:  

1. `x`: Displacement in the horizontal direction as a function of time (in meters) 
2. `vx`: Speed in the horizontal direction as a function of time (in m/s units) 
3. `y`: Displacement in the vertical direction as a function of time (in meters) 
4. `vy`: Speed in the vertical direction as a function of time (in m/s units) 
5. `y_x`: Displacement in the vertical direction as a function of horizontal displacement (in meters) 

In all cases, it is assumed that there are no motion in the lateral direction.

### Projectile Motion in Vaccum  
Arguments of the `projectile` function are:  

1. `v0`: Initial velocity in m/s 
2. `y0`: Initial height in m 
3. `theta0`: Initial angle in degrees 
4. `t`: Time of flight in seconds 

For vectors of length `1` for all arguments, the function will return a list of
projectile parameters for those arguments. For example for an initial velocity `v` of
`30 m/s`, initial height `y` of `0 m`, initial projectile angle `theta0` of `30 degrees`, and 
at time `t=3 seconds` we get:


```{r}
projectile(t = 3, y0 = 0, v0 = 30, theta0 = 30)
```

One can also pass a vector of length `> 1` for any individual input parameter. This is 
most useful for the time parameter `t`:

```{r}
vacuum <- projectile(t = seq(0,5,0.1), y0 = 0, v0 = 50, theta0 = 30)
plot(vacuum$x, vacuum$y, pch=19, col="blue", main="Projectile motion in vacuum")
```


```{r}
t <- seq(0,5,0.1)
vacuum <- projectile(t = t, y0 = 0, v0 = 50, theta0 = 30)
## prepare grid
par(mfcol=c(2,2), mar = c(3.5,3,1,1), oma=c(2,2,2,2), mgp=c(2.2,1,0))
## plot outputs
plot(t, vacuum$x, pch=19, col="blue")
plot(t, rep(vacuum$vx, length(t)), pch=19, col="blue")
plot(t, vacuum$y, pch=19, col="red")
plot(t, vacuum$vy, pch=19, col="red")
title("Projectile motion in vacuum", outer=TRUE)
```


### Projectile Motion in Air  
Arguments of the `projFrictionLin` function are:  

1. `v0`: Initial velocity in m/s 
2. `y0`: Initial height in m 
3. `theta0`: Initial angle in degrees 
4. `t`: Time of flight in seconds
5. `b`: drag coefficient in Newtons.seconds/meters
6. `m` mass of object in kg

For vectors of length `1` for all arguments, the function will return a list of
projectile parameters for those arguments. For example for an initial velocity `v` of
`30 m/s`, initial height `y` of `0 m`, initial projectile angle `theta0` of `30 degrees`, 
drag coeffiecient `b` of 0.5, mass of projectile `m` of 5 kg,  and 
at time `t=3 seconds` we get:


```{r}
projFrictoinLin(t = 3, y0 = 0, v0 = 30, theta0 = 30, b = 0.5, m = 5)
```

One can also pass a vector of length `> 1` for any individual input parameter. This is 
most useful for the time parameter `t`:

```{r}
friction <- projFrictoinLin(t = seq(0,5,0.1), y0 = 0, v0 = 50, theta0 = 30, b = 0.5, m = 5)
plot(friction$x, friction$y, pch=19, col="blue", main="Projectile motion in Air")
```


```{r}
t <- seq(0,5,0.1)
friction <- projFrictoinLin(t = t, y0 = 0, v0 = 50, theta0 = 30, b = 0.5, m = 5)
## prepare grid
par(mfcol=c(2,2), mar = c(3.5,3,1,1), oma=c(2,2,2,2), mgp=c(2.2,1,0))
## plot outputs
plot(t, friction$x, pch=19, col="blue")
plot(t, friction$vx, pch=19, col="blue")
plot(t, friction$y, pch=19, col="red")
plot(t, friction$vy, pch=19, col="red")
title("Projectile motion in Air", outer=TRUE)
```


### Comparing Projectile Motion   

In the example below we compare projectile motion in vacuum to that in air:  


```{r, tidy.opts=list(width.cutoff=19)}
par(mfcol=c(1,1))
vacuum   <- projectile(t = seq(0,5,0.1), y0 = 10, v0 = 50, theta0 = 30)
friction <- projFrictoinLin(t = seq(0,5,0.1), y0 = 10, v0 = 50, theta0 = 30, b = 0.5, m = 5)

x  <- vacuum$x; y  <- vacuum$y; xf <- friction$x; yf <- friction$y_x

plot(x, y, col="blue", pch=19, ylim=c(min(y, yf), max(y, yf)),
     xlab = "Horizontal Distance Travelled (m)",
     ylab = "Vertical Distance Travelled (m)")
points(xf, yf, col="red", pch=21)
legend(x="topleft", legend = c("No Friction", "Friction"),
    col = c("blue", "red"), pch = c(19, 21))
```

## The jit Function
The `jit` function is used to randmoize a given vector. It has two modes, `norm` and `uniform`:  

*`norm`: randomizes a numerical vector by drawing samples from a noraml distribution.
*`uniform`: randomizes a numerical vector by drawing samples from a uniform distribution.  

In the example below we randomize the number `40`, `5` times by drawing from a normal distribution 
with mean equal to the value to be ranomized, `40`, and a standard deviation equal to 3, passed to 
the `mean` argument:

```{r}
jit(x = 40, n = 5, method = "norm", amount=3)
```

The example below does a similar ranomization, jittering, but with a uniform distribution:

```{r}
jit(x=40, n=5, method="uniform")
```

Please notice that the `jit` function with the `uniform` method is essentially equivalent to the `jitter` function in R. 

```{r, tidy.opts=list(width.cutoff=10), fig.height=6}
par(mfrow=c(2,1), mar = c(3.5,3,1,1), oma=c(2,2,2,2), mgp=c(2.2,1,0))
hist(jit(x=1, n=1000, method="norm", amount = 1), main="Normal Dist.", xlab="Randomized Values", col="lightblue")
hist(jit(x=1, n=1000, method="uniform", amount = 1), main="Uniform Dist.", xlab="Randomized Values", col="lightblue")
```

# Simulations

## Simulation Parameters:
1. $\theta$: Angle with respect to the vertical direction. 
2. $\phi$: Lateral angle, angle with respect to the horizontal direction.
3. $t$: Time of firing.

We will be simulating the following uncertainities:

1. $\delta\theta$: Uncertainity in $\theta$
2. $\delta\phi$: Uncertainity in $\phi$
3. $\delta t$: Uncertainity in firing time $t$

A detailed example of generating simulations can be found in `example_2` in the `example` directory. An output of this exmaple is shown below:

```{r, echo=FALSE}
## initialize parameters
t      <- seq(0,5,0.1)
v0     <- 50
y0     <- 0
theta0 <- 30
b      <- 0.5
m      <- 5

## We will assume the truth to be the projectile motion as initialized
## by the values above
set.seed(pi)
truth <- projFrictoinLin(t = t, y0 = y0, v0 = v0, theta0 = theta0, b = b, m = m)

## Simulate n number of runs while slightley changing each input parameter
Sim <- function(t = seq(0,5,0.1), y0 = 0, v0 = 50, theta0 = 30, b = 0.5, m = 5, 
                n = 5, amount=1, seed=NULL){
        #function to simulate projectile motion
        set.seed(seed)
        df <- data.frame(y0=jit(y0,n, amount), 
                         v0=jit(v0,n, amount), 
                         theta0=jit(theta0,n, amount),
                         b = jit(b,n, amount),
                         m=jit(m,n, amount))       
        ## List of projectile data for MC runs
        MC <- list()     
        for(i in 1:n){
                MC[[i]] <- with(df, projFrictoinLin(t = t, y0 = y0[i], 
                                                    v0 = v0[i], 
                                                    theta0 = theta0[i], 
                                                    b = abs(b[i]), 
                                                    m = abs(m[i])))
        }
        # return list
        return(MC)
}

makePlot <- function(simList=NULL, par=2, N){
        ## function to plot projectile parameter as a function to time
        if(par==1){
                parplot="x"
                yLab = "Horizontal Disp. (m)"
        }
        else if(par==2){
                parplot="y"
                yLab="Vertical Disp. (m)"
        }
        if(par==3){
                parplot="vx"
                yLab = "Horizontal Velocity (m)"
        }
        else if(par==4){
                parplot="vy"
                yLab="Vertical Velocity (m)"
        }
        Min <- min(simList[[1]][[parplot]])
        Max <- max(simList[[1]][[parplot]])
        if(parplot=="vx" | parplot=="vy"){loc="bottomleft"}
        else{loc="topleft"}
        for(i in 1:N){
                Min <- min(Min, simList[[i]][[parplot]])
                Max <- max(Max, simList[[i]][[parplot]])
        }
        
        plot(t, truth[[parplot]], ylim=c(Min, Max), pch=19, col="blue",
             ylab=yLab, xlab="Time (s)")
        for (i in 1:N){
                points(t, simList[[i]][[parplot]], col=i, pch=1)
        }
        legend(x=loc, legend = c("Truth", "MC"), pch=c(19, 1),
               col=c("blue", 1))
}

## Make some plots
## prepare the gird layout
par(mfcol=c(2,2), mar = c(3.5,3,1,1), oma=c(2,2,2,2), mgp=c(2.2,1,0))

## we'll simulate 5 runs. To change the number of simulated runs
## just change N below. Very large N's will result in overcrowded plots
N <- 5
## set seed below for reproducibility purposes
seed = 123
## set seed to NULL for random seed
# seed = NULL

makePlot(simList = Sim(n = N, amount = 1, seed = seed), par = 1, N = N)
makePlot(simList = Sim(n = N, amount = 1, seed = seed), par = 2, N = N)
makePlot(simList = Sim(n = N, amount = 1, seed = seed), par = 3, N = N)
makePlot(simList = Sim(n = N, amount = 1, seed = seed), par = 4, N = N)
title("Simulating Projectile Motion", outer=TRUE)
```