---
title: "VVnA R Package"
author: "Aous Abdo"
date: "Tuesday, August 12, 2014"
opset: pdf
output:
  pdf_document:
    toc: TRUE
  knitrBootstrap::bootstrap_document:
    title: "VVnA R Package"
    theme: Simplex
    highlight: Solarized - Light
    theme.chooser: FALSE
    highlight.chooser: FALSE
    menu: TRUE
---
```{r echo=FALSE}
library(knitr)
library(VVnA)
```
# Introduction
 
The VVnA "Validation, Verification, and Accreditation" package is a package intended 
for ....

# Functions
## Projectile Motion
Projectile motion in vacuum and in air are calculated with the `projectile` and `projFrictionLin` functions respectively. when considering air friction effects on projectiles, we only consider the viscous drag which is related to the 
velocity `v`. The Inertial drag realted to the square of the velocity is not treated in this package. 

In each of these two cases, a function will return the following projectile parameters:  

1. `x`: Displacement in the horizontal direction as a function of time (in meters)  
2. `vx`: Speed in the horizontal direction as a function of time (in m/s units)  
3. `y`: Displacement in the vertical direction as a function of time (in meters)  
4. `vy`: Speed in the vertical direction as a function of time (in m/s units)  
5. `y_x`: Displacement in the vertical direction as a function of horizontal displacement (in meters)  

In all cases, it is assumed that there are no motion in the lateral direction.

### Projectile Motion in Vaccum  
Arguments of the `projectile` function are:  

1. `v0`: Initial velocity in m/s  
2. `y0`: Initial height in m  
3. `theta0`: Initial angle in degrees  
4. `t`: Time of flight in seconds  

For vectors of length `1` for all arguments, the function will return a list of
projectile parameters for those arguments. For example for an initial velocity `v` of
`30 m/s`, initial height `y` of `0 m`, initial projectile angle `theta0` of `30 degrees`, and 
at time `t=3 seconds` we get:


```{r}
projectile(t = 3, y0 = 0, v0 = 30, theta0 = 30)
```

One can also pass a vector of length `> 1` for any individual input parameter. This is 
most useful for the time parameter `t`:

```{r}
vacuum <- projectile(t = seq(0,5,0.1), y0 = 0, v0 = 50, theta0 = 30)
plot(vacuum$x, vacuum$y, pch=19, col="blue")
```

```{r}
t <- seq(0,5,0.1)
vacuum <- projectile(t = t, y0 = 0, v0 = 50, theta0 = 30)
## prepare grid
par(mfcol=c(2,2), mar = c(3.5,3,1,1), oma=c(2,2,2,2), mgp=c(2.2,1,0))
plot(t, vacuum$x, pch=19, col="blue")
points(t, rep(vacuum$vx, length(t)), pch=19, col="blue")
points(t, vacuum$y, pch=19, col="blue")
```
